\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-11-04 R/clean_spectral_data.Rmd
updated some wording to reflect species vs genus
since the genus taxonRanks have a species of sp, they needed to be exluded from the initial count of unique species values. and I changed the wording later from species to "species and genus" or "scientific name label", recognizing that one of the classes was Larix sp, which is a genus taxon rank

message("# of unique species values:")
length(unique(all_spectra$species))

@@

```{r}
# Histogram with just the taxons that have enough samples 
samplesCountGT <- taxonCounts %>% 
  filter(countGT == TRUE)

>>>>

# not including spectra with "sp" as the species, when taxonRank is genus
message("# of unique species values:") 
tmp <- all_spectra %>% filter(taxonRank != "genus") %>% select(species, taxonRank)
length(unique(tmp$species))

@@

```{r fig.width=9, fig.height=12}
# Histogram with just the taxons that have enough samples 
samplesCountGT <- taxonCounts %>% 
  filter(countGT == TRUE)

### 2020-11-04 analysis/paper/paper.Rmd
updated some wording to reflect species vs genus
since the genus taxonRanks have a species of sp, they needed to be exluded from the initial count of unique species values. and I changed the wording later from species to "species and genus" or "scientific name label", recognizing that one of the classes was Larix sp, which is a genus taxon rank

We performed our analysis in R [@R] with packages including neonUtilities [@neonUtilities], geoNEON [@geoNEON], neonhs [@neonhs].
In June 2020, we extracted hyperspectral reflectance data for all individual mapped stems at NEON sites that had an AOP flyover during the same year as when *in-situ* measurements were collected. 
We only kept individuals with a "live" plant status (as opposed to other plant status classifications such as "dead", "standing dead", "downed", "insect/disease damaged", and "no longer qualifies").
This included spectral reflectance spectra for 5972 observations of individual plants. 
Some individuals had multiple observations across years. 
The individual plants were identified taxon ranks of kingdom, family, genus, species, and variety. 
Most of them were identified at the species rank.
We filtered the plants by taxonomic rank to retain only those identified at the genus, species, and variety ranks. 
We removed varieties from scientific names to label each individual spectrum with its respective genus and species (when available). 
There were 74 unique genus values, 118 unique species values, and 151 unique genus-species combinations.
The hyperspectral reflectance spectra extracted from the locations of the individual plants have different starting wavelengths:

@@

Next, we removed samples belonging to species classes with few samples. 
We figured that a small sample size would lead to a poor representation in model training. 
From the initial collection of 146 species, 50 species had at least 20 samples (Figure \@ref(fig:fig-histSpecies)). 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 species (and therefore, all genus values) were represented in each set.

@@

**Figure - histogram of species counts? maybe in appendix, make a table or something to show species info for which ones were used after cleaning/filtering**
**Figure - cleaned data, make it legible, how to color if at all (the species/genus/variety color doesn't make sense anymore at this stage)...**

@@

It ain't much...

>>>>

We performed our analysis in R [@R] with packages including neonUtilities [@neonUtilities], geoNEON [@geoNEON], neonhs [@neonhs].
In June 2020, we extracted hyperspectral reflectance data for all individual mapped stems at NEON sites that had an AOP flyover during the same year as when *in-situ* measurements were collected. 
We only kept individuals with a "live" plant status (as opposed to other plant status classifications such as "dead", "standing dead", "downed", "insect/disease damaged", and "no longer qualifies").
This included spectral reflectance spectra for 5972 observations of individual plants. 
Some individuals had multiple observations across years. 
The individual plants were identified taxon ranks of kingdom, family, genus, species, and variety. 
Most of them were identified at the species rank.
We filtered the plants by taxonomic rank to retain only those identified at the genus, species, and variety ranks. 
We removed varieties from scientific names to label each individual spectrum with its respective genus and species (when available). 
There were 74 unique genus values, 113 unique species values, and 151 unique genus-species combinations.
The hyperspectral reflectance spectra extracted from the locations of the individual plants have different starting wavelengths:

@@

Next, we removed samples belonging to species and genus classes with few samples. 
We figured that a small sample size would lead to a poor representation in model training. 
From the initial collection of 151 unique scientific names identified at the genus or species level, 50 of them had at least 20 samples and were kept for the classification analysis (Figure \@ref(fig:fig-histSpecies)). 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 scientific name labels were represented in each set.

@@

**Maybe in appendix, make a table or something to show species info for which ones were used after cleaning/filtering**

**Figure - cleaned data, make it legible, how to color if at all (the species/genus/variety color doesn't make sense anymore at this stage)...**

@@

Histogram showing which scientific names have at least 20 samples to be included in the classification analysis. Note that the "Larix sp." invididuals are classified at the genus level.
